\chapter*{Acknowledgments}
%\addcontentsline{toc}{chapter}{Acknowledgments}


%Mentor for life: career development and long-term interest
    %2.    Enthusiasm: for science, for the student's project and the student
    %3.    Sensitivity: to personal and professional needs and circumstances
    %4.    Appreciating individual differences
    %5.    Respect
    %6.    Unselfishness: "lack of intellectual jealousy"
    %7.    Supports others: outside their own sphere of responsibility
    %8.    Teaching and communication skills
		
		%
		%Supervisors must:
%
%ensure you get appropriate credit for your work
%ensure your research is valid and accurate

I am sincerely grateful to the many people who have contributed in some way to the completion of this thesis.
My PhD supervisory panel and collaborators, Sylvie Thi{\'e}baux, Menkes van den Briel, Hassan Hijazi, Christfried Webers and Pascal Van Hentenryck have been extremely supportive during these past four years. 

Pascal pointed us to optimisation problems in the smart building space, seeding the topic of this thesis.
Working on a cross-diciplinary research topic is never easy. Even more so when I started with nearly zero background in HVAC and optimisation. 
Sylvie has provided invaluable insights and immense guidance on my research. 
She has given me the freedom to explore various research topics, pushed me to test my limits and connected me to the right people whenever I stumbled into obstacles. 
Meanwhile, she knew when to pull me back on track and help me stay focus throughout each milestone. 
I enjoyed our time spent together writing papers and she was instrumental in helping me crank out the papers that we published together, one in particular with only a few hours to spare before the deadline!
I love our conversations and her company throughout the regular discussion meetings and our trips together to AAAI'15, CP'16, CSIRO Newcastle etc.
Beyond research, I greatly appreciate her close personal support over the years. She taught me to treat life from different perspectives, from lifestyle, food, music to relationships, which significantly differ due to the distinct cultural differences between us. 
I think the decision of writing an email to her for a PhD position back in May 2012 might be one of the most life-changing decisions I have made.
She was and remains my best role model as a scientist, mentor and teacher.

Menkes sparked my interest on mixed integer programming. 
He taught me to spend more time on understanding a problem, instead of formulating a solution quickly with an existing model at hand. 
He inspired me to create simple models that work efficiently. He taught me that all models are wrong, hence developing one that is useful, is more essential than solving all problems at once in a single model.
I'd definitely have a rough time without him in my supervisory panel, simply because I feel comfortable debugging a model with him and asking him silly questions, and then things work out pretty well! 
The office atmosphere has changed ever since his departure to start a company. I still think fondly of the funny sounds and finger gestures he made to cheer-up everyone on the floor. 

Despite him not being on my supervisory panel, I have worked closely with Hassan on non-linear programming and robust optimisation. He struck me as a passionate researcher when he came to me a day before the conference deadline, and said to me:``Hey Boon Ping, I've a better idea to model the solution!''. It was impossible to re-run large-scale experiments within a day, but indeed his attitude influences me to keep on thinking and improving on the solution without being limited by a deadline. 

Whilst I did not end up working on a Machine Learning topic, Chris has been helpful in providing career advice and suggestions many times during my PhD study. I enjoyed very much the long chats we had periodically after each milestone.

I am indebted to all of them because without their help this thesis would never have been completed.

I like to thank all current and past students/co-researchers in Data61/CSIRO, formerly NICTA, particularly those that I have had the pleasure of spending my time with: Paul Scott, Terrence Mak, Buyu Liu, Jing Cui, Ksenia Bestuzheva, Kar-Wai Lim, Swapnil Mishra, Arthur Maheo, Cody Christopher, Franc Ivankovic, Frank Su, Fazlul Hasan Siddiqui, Karsten Lehmann, Alan Lee and Felipe Maldonado Caro. Thanks to Jun-Yong Kwak for generously sharing his research data. A special word of thanks goes to Phil Kilby, Patrik Haslum, Alban Grastien, Andreas Schutt, Tommaso Urli, Marian-Andrei Rizoiu, Enrico Scala, Dan Gordon, Carletton Coffrin, Michael Norrish, Lexing Xie, MiaoMiao Liu and Cheng-Soon Ong. Their generous sharings on technical knowledge in optimisation, planning and scheduling, large neighbourhood search, constraint programming, energy systems, machine learning and research skills in general have been invaluable. Furthermore, I would miss all the great fun we had playing foosball, literally every day, and board games every Monday. I still get a laugh when I reminisce about us confusing, tricking and bluffing each other while playing Avalon. After working for 7 years, it was a tough decision opting out of climbing the corporate ladder and to follow my childhood dream. However, I have certainly appreciated and enjoyed doing research throughout my stay at ANU/Data61. Each person listed has played some role in my time here. 

Additionally, I'd like to thank the Malaysian community in Canberra, especially Ayeli de Marillac and Yee-Harn Teh, for making my stay here so much more enjoyable and homely. I also thank my friends back home, especially Ligin Lem, Swee-Kean Chiam, Siew-Kiang Tan, Sok-Hooi Teng, Shiek-Wei Tan, Sheau-Horng Ng, Zhu Ying, Pitt Ong and Chin-Chee Cheah, for providing the support and friendship that I needed. We were in touch, literally, 24/7 through whatsapp, messenger, telegram, wechat and LINE. These past several years have not been an easy ride, both academically and personally. Without their constant encouragement, I would not have gone this far.  

I would also like to thank Ettikan K. Karuppiah, Eiichi Muramoto, Poh-Kit Chong and Chia-Ching Ooi for playing an important role during the application stage of my PhD position. They wrote me reference letters, reviewed my statement of purpose and advised me about how to survive in a PhD program. Over the years I was comfortably settling in my corporate job, Chia-Ching and Poh-Kit were constantly nagging me about when I would pursue a PhD. It is great that I have eventually fulfilled my word after more than 10 years!

I especially thank my dad, mum, brothers, sister and our extended family for their understanding and unconditional support.
I know I always have my family to count on when times are rough. 
A special word of thanks goes to my younger brother, for showing us miracles after a dreadful accident that almost took his life in November 2016. 
He has once again shown to us that being adventurous, persistent and never giving up are our common family traits. 
That holds us through ups and downs, and drives us to live life to the fullest. 
